How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-17 tse-paper/content.tex
Formatting

Continuous Integration (CI) has become a common practice in software
engineering~\cite{hilton2016usage}.
Many software projects use
CI~\cite{hilton2016usage,staahl2014modeling,beller2017oops} to detect
bugs early~\cite{vasilescu2015quality,duvall2007continuous}, improve
developer productivity~\cite{miller2008hundred,hilton2016usage} and
communication~\cite{downs2012ambient}.
CI builds produce logs which
report results of various sub-steps within the build.
These build logs
contain a lot of valuable information for developers and researchers---for
example, descriptions of compile errors, linter warnings or failed
tests~\cite{beller2017oops,seo2014programmers,vassallo2017a-tale}.

CI builds produce logs which report the results of the steps within
the build.
The information stored in build logs can and has
already been used for a variety of applications.
One prominent dataset
in the space, TravisTorrent, was featured as the MSR Data Challenge
2017~\cite{msr17challenge}.
To date, scientist and practitioners alike
have used it and other, proprietary datasets to understand and
cateogrize Continuous Integration build
failures~\cite{islam2017insights}, to do research on testing
practices~\cite{orellana2017differences}, to train classifiers to
predict the build outcome and
duration~\cite{ni2017cost,bisong2017built,machalica2019predictive},
and to investigate static analysis tools in CI
builds~\cite{zampetti2017open}.
Therefore, being able to efficiently
and correctly extract information from build logs is paramount to the
future of a variety of fields depending on it.

However, build logs can be verbose and large---sometimes in excess of
50 MB of ASCII text~\cite{beller2017oops}---making them inadequate
for direct human consumption.
Therefore, to support developers and
researchers in efficiently making use of the information within build
logs, we must at least semi-automatically retrieve the chunks of the
log that describe the targeted information.

There are different techniques to retrieve information chunks from CI
build logs.
Beller et al.\ use a rule-based system of regular
expressions to analyze logs from Travis CI~\cite{beller2017oops}.
Such regular expressions are developed by looking at exemplary build
logs.
Vassallo et al.\ wrote a custom parser to gather information
for build repair hints~\cite{vassallo2018un-break}.
Recently, Amar et
al.\ reduced the number of lines for a developer to inspect by
creating a diff between logs from failed and successful
builds~\cite{amar2019mining}.

These approaches have various strengths and weaknesses: Regular
expressions are exact, but tedious and error-prone to
maintain~\cite{michael2019regexes}.
Custom parsers are powerful
though fragile in light of changes in the log structure.
Diffing
between failed and successful logs can reduce the information to be
processed, but is at best semi-automatic~\cite{amar2019mining}.

% TODO: 4 sentence description of the steps in graphic?
\Cref{fig:overview} gives an overview of our research steps.
We conducted a systematic mapping study,
to gain an impression on how researchers are using information from
build logs within their research or to support
developers.
% We implemented three promising techniques that retrieve
% chunks of text from build logs and can be adapted to
% a specific project and use case by giving examples of build logs and
% correpsonding chunks.
% We evaluated the
Our goal was to clarify the state-of-research on build log
analysis (\textbf{RQ1}).

The outcomes of our study shows that there are various attempts to
retrieve specific information from buildlogs.
The resulting implementations are rarely available and therefore rarely
reused.
Because of the high development overhead
of the custom parsers, many implementations are limited to small number
of supported build tools.

\begin{figure*}[htb]
	\centering
	\includegraphics[width=\textwidth, trim={1.2cm 10.5cm 1.2cm 0cm},
	clip]{img/overview.pdf}
	\caption{Our Research Workflow.}
	\label{fig:overview}
\end{figure*}

To address this, we implement three promising techniques for build
log analysis.
They retrieve specific chunks of targeted information and are trained
with examples provided by the user.
We conduct an empirical study on the \emph{LogChunks} data
set~\cite{brandt2020logchunks} to
gauge the performance of these chunk retrieval techniques and characterize
when one of the techniques should be preferred over other alternatives
(\textbf{RQ2}).

\begin{simplebox}[minipage boxed title*=-5cm]{Research Questions}
\begin{itemize}
  \item[\textbf{RQ1:}] What is the state-of-research on build
  log analysis?
  \item[\textbf{RQ2:}] When are PBE, CTS or KWS suited to
  retrieve chunks from build logs?
\end{itemize}
\end{simplebox}

% also moved down to where mentioned in detail
% left copy here for convenience

% \begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
% {\textbf{RQ1:} What is the state-of-research on build
% log analysis?}
% \begin{itemize}
%   \item[\textbf{RQ1.1:}] Which information is targeted?
%   \item[\textbf{RQ1.2:}] Which techniques are used?
%   \item[\textbf{RQ1.3:}] Which build logs are analyzed?
% \end{itemize}
% \end{simplebox}

% \begin{simplebox}[minipage boxed title*=-2.5cm,
% attach boxed title to top center={yshift=-6mm}]
% {\textbf{RQ2:} When are PBE{,} CTS or KWS suited to
% retrieve chunks from build logs?}
% \begin{itemize}
%   \item[\textbf{RQ2.1:}] How many examples does a technique need to
%   perform best?
%   \item[\textbf{RQ2.2:}] How structurally similar do the examples
%   need for a technique to be applicable?
%   \item[\textbf{RQ2.3:}] How accurate are the retrievals of a technique?
% \end{itemize}
% \end{simplebox}

% % At the moment there is only anecdotal evidence on the performance
% of these
% % techniques and on when a technique should be preferred over other
% alternatives.
% % Developers and researches currently have little support when choosing
% which
% % technique to use for a task.

% % The goal of this article is to investigate different chunk retrieval
% techniques
% % for build logs and describe under which circumstances certain
% techniques
% can be
% % recommended over others.
% % We aim to characterize different chunk retrieval techniques.
% % For \textbf{Research Question 1} (\textbf{RQ1}), we analyze which
% criteria
% % influence the suitability of a chunk retrieval technique for CI
% build logs.

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on user examples of build logs and chunks, PBE
  synthesizes
  a regular expression matching the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  Using the Vector Space Model, CTS selects those lines of a log which are
  most similar to the lines present in example chunks provided by
  the user.
  \item \textbf{Keyword Search (KWS)}
  Ad-hoc method for finding fitting passages by searching the whole
  text for
  the occurrence of specific trigger words.
\end{itemize}

Following the results from our comparison study,
recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in precision
and recall of the chunk retrieval is required.
Results produced by PBE are suited for automatic on-ward processing.
CTS is well suited when the representation of the desired information
varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired information in
the log
is unpredictable or varies greatly, KWS seems to be the best choice.
However,
its low precision---it extracts a context of multiple lines around a
finding---makes it generally unsuited for automatic on-ward processing and
instead requires a human to further inspect and interpret the output
chunk.

>>>>

Within the past two decades, Continuous Integration (CI) has become a
ubiquitous best practice to streamline the build process of software
projects~\cite{hilton2016usage,beller2017oops,staahl2014modeling}.
Build
logs are a textual by-product that automatic software build processes
create.
As a ``treasure trove'' of information, they contain results
not only of the elemntary steps to ``make'' a project---such as
retrieving and resolving dependencies, compiling, and testing the
software---but also about anciliary quality assurance steps such as
automated, static analysis tools.

When a build or one of its sub steps fails, developers typically
scavenge the log for the source of the error.
This manual activity is
tedious and prone to error~\cite{santolucito2018statically}.
As a
single build log can be over 50 megabytes large~\cite{beller2017oops},
it finding a piece of information in it is akin to searching for a
needle in a haystack.
In addition, due to their information richness,
build logs enable a plethora of other applications: with the help of
build logs, we can better understand and cateogrize Continuous
Integration failures and compile
errors~\cite{islam2017insights,seo2014programmers}, differentiate
testing practices~\cite{orellana2017differences,vassallo2017a-tale},
train classifiers to predict the build outcome and
duration~\cite{ni2017cost,bisong2017built,machalica2019predictive}, or
investigate static analysis tools in CI
builds~\cite{zampetti2017open}.
Manual analysis of build logs does not
scale to the number of build logs and amount of information that such
sophisticated applications require.
Therefore, being able to
automatically extract pieces of information from build logs
is not only paramount to such applications, but can also help
developers debug a broken build.

However, despite its paramount role, the process of how to analyze
build logs has so far received little attention, bringing us to our
first research question:
\begin{simplebox}[minipage boxed title*=-5cm]{Research Question}

\begin{itemize}
  \item[\textbf{RQ1:}] What are the state-of-the-art techniques on build
  log analysis?
 \end{itemize}
\end{simplebox}

In this article, as first step (\Cref{fig:overview}), we show that
less than half of the 61 works which rely on build log analysis
describe how they performed it, and even fewer did so in detail, let
alone made their implementation available.
Even so, the literature
survey finds three main strategies to extract information chunks from
build logs: Using 1) a set of manually curated regular expressions,
similar to TravisTorrent~\cite{beller2017oops}, 2) a list of keywords
to identify regions of interest in the build log, and 3)
% TODO (Caro): Provide better description
doing text-similarity.

Even though these three approaches to retrieve information chunks from
build logs have different strengths and weaknesses, the literature
survey shows that their use is often arbitrary and that there is
usually no discussion on why to use one over the other.
Developers and
researchers not only seem to lack awareness of which techniques exist
(RQ1), but also guidelines on when to employ which.
As a consequence,
in RQ2 we ask:

\begin{simplebox}[minipage boxed title*=-5cm]{Research Question}
\begin{itemize}
  \item[\textbf{RQ2:}] What are the comparative strengths and weaknesses
  of the top-3 build log anaylsis techniques?
\end{itemize}
\end{simplebox}

To answer this question, we implement three automatic techniques for
build log analysis based on the three stratgies at large (step 2):

\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on user examples of build logs and chunks, PBE
  synthesizes
  a regular expression matching the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  Using the Vector Space Model, CTS selects those lines of a log which are
  most similar to the lines present in example chunks provided by
  the user.
  \item \textbf{Keyword Search (KWS)}
  Ad-hoc method for finding fitting passages by searching the whole
  text for
  the occurrence of specific trigger words.
\end{itemize}

In step 3, we conduct an empirical study to gauge the performance of
these chunk retrieval techniques under a variety of conditions in
real-world projects.
Finally, we develop guidelines (step 4) on when
one of the techniques might be perferrential to its alternatives.

\begin{figure*}[htb]
	\centering
	\includegraphics[width=\textwidth, trim={1.2cm 10.5cm 1.2cm 0cm},
	clip]{img/overview.pdf}
	\caption{Research Design.}
	\label{fig:overview}
\end{figure*}

We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in
precision and recall of the chunk retrieval is required.
Results
produced by PBE are suited for automatic on-ward processing.
CTS is
well suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired
information in the log is unpredictable or varies greatly, KWS seems
to be the best choice.
However, its low precision---it extracts a
context of multiple lines around a finding---makes it generally
unsuited for automatic on-ward processing and instead requires a human
to further inspect and interpret the output chunk.

### 2020-04-17 tse-paper/paper.tex
correct first half of lit survey
add corrections ttv - conclusion

\newcommand{\smallish}{\fontsize{6.15pt}{20pt}\selectfont}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Carolin~Brandt,~\IEEEmembership{Member,~IEEE,}
        Moritz~Beller,~\IEEEmembership{Fellow,~OSA,}
        and~Annibale~Panichella,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell was with the Department
of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
GA, 30332.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: see http://www.michaelshell.org/contact.html
\IEEEcompsocthanksitem J. Doe and J. Doe are with Anonymous University.}% <-this % stops an unwanted space
\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

@@

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Computer Society, IEEE, IEEEtran, journal, continuous integration, build log analysis, chunk retrieval.
\end{IEEEkeywords}}

@@

% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

The authors would like to thank...

@@

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

@@

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

>>>>

\newcommand{\smallish}{\fontsize{6.15pt}{20pt}\selectfont}
\newcommand{\tinyish}{\fontsize{6.5pt}{6.15pt}\selectfont}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Carolin~Brandt,~\IEEEmembership{Member,~IEEE,}
        Moritz~Beller,~\IEEEmembership{Fellow,~OSA,}% <-this % stops a space
% \IEEEcompsocitemizethanks{\IEEEcompsocthanksitem M. Shell was with the Department
% of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
% GA, 30332.\protect\\
% % note need leading \protect in front of \\ to get a newline within \thanks as
% % \\ is fragile and will error, could use \hfil\break instead.
% E-mail: see http://www.michaelshell.org/contact.html
% \IEEEcompsocthanksitem J. Doe and J. Doe are with Anonymous University.}% <-this % stops an unwanted space
\thanks{Manuscript received May 1, 2020; revised August 26, 2015.}}

@@

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Computer Society, IEEE, IEEEtran, Journal, Continuous Integration, 
Build Log Analysis, Chunk Retrieval.
\end{IEEEkeywords}}

@@

% use section* for acknowledgment
% \ifCLASSOPTIONcompsoc
%   % The Computer Society usually uses the plural form
%   \section*{Acknowledgments}
% \else
%   % regular IEEE prefers the singular form
%   \section*{Acknowledgment}
% \fi

% The authors would like to thank...

@@

\begin{IEEEbiography}{Carolin E. Brandt}
Biography text here.
\end{IEEEbiography}

\begin{IEEEbiography}{Moritz Beller}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
% \begin{IEEEbiographynophoto}{John Doe}
% Biography text here.
% \end{IEEEbiographynophoto}

@@

% \begin{IEEEbiographynophoto}{Jane Doe}
% Biography text here.
% \end{IEEEbiographynophoto}
